---
title: "Analysis of Activity Monitoring Data"
author: "Farhan Atiq"
date: "July 29, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

##Load in the data and required libraries

This analysis requires the ggplot2 and lubridate packages

```{r}
library(ggplot2)
library(lubridate)
```

```{r}
data <- read.csv("C:/Users/fatiq/Documents/R/Coursera/activity.csv")

```

## What is mean total number of steps taken per day?

For this part of the assignment, you can ignore the missing values in the dataset

```{r}

steps.by.date <- aggregate(steps ~ date, data = data, FUN = sum)
ggplot(data = steps.by.date, aes(steps)) + 
  geom_histogram(bins = 30) + 
  xlab("Steps") + 
  ylab("Count")

``` 

Mean and Median steps per day
```{r}

mean(steps.by.date$steps)
median(steps.by.date$steps)

```

#What is the average daily activity pattern?

Average steps per 5-minute interval

```{r}
steps.by.time <- aggregate(steps ~ interval, data = data, FUN = mean)
ggplot(data = steps.by.time, aes(interval, steps)) + 
  geom_line() + 
  xlab("Interval") + 
  ylab("Steps")

```


Highest Activity Interval

```{r}
steps.by.time[which.max(steps.by.time$steps),]$interval
```

The 5-minute interval that contains the maximum of steps, on average across all days, is 835.


#Imputing Missing Data

First copy the original data set

```{r}

data.imputed <- data

```

Change all Na's to the mean value (excluding Na's)

```{r}

mean.steps <- mean(data$steps, na.rm = TRUE)
data.imputed[is.na(data.imputed)] <- mean.steps
```

Histogram of steps per day including imputed values

```{r}

imputed.by.date <- aggregate(steps ~ date, data = data.imputed, FUN = sum)
ggplot(data = imputed.by.date, aes(steps)) + 
  geom_histogram(bins = 30) + 
  xlab("Steps") + 
  ylab("Count")
```


#Are there differences in activity patterns between weekdays and weekends?

Determine which days are weekends and include that in aggregation

```{r}
isWeekday <- function(x) {
  day <- wday(x)
  return(day %in% seq(1,5))
}
data$is.weekday <- isWeekday(data$date)
agr <- aggregate(steps ~ interval + is.weekday, data = data, FUN = mean)
agr$is.weekday <- as.factor(agr$is.weekday)
levels(agr$is.weekday) <- c('Weekday','Weekend')
```
Plot activity

```{r}
ggplot(data = agr, aes(interval, steps)) + 
  geom_line() + 
  xlab("Interval") + 
  ylab("Steps") + 
  facet_grid(is.weekday ~ .)
```
